Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-06-09 paper/letter.tex
new multiphase info

\tol\ is a compact star forming dwarf galaxy that presents a
strong Ly-$\alpha$ emission \cite{Thuan97} with two important
features. 
First, it has symmetric profile  without any systemic displacement
from the expected line-center. 
Second, it shows a triple peaked profile. 
Commonly, the Ly-$\alpha$ line features a single or asymmetric double peak. 
These two special features in \tol\ had evaded a natural physical
interpretation so far.

We find that a rotating spherical gas distribution provides a
satisfactory explanation of \tol's puzzling features.  
Figure 1. summarizes our findings.
Dots represent the observational data for \tol with the
overplot from our best fit model from the full radiative transfer
simulation.   
The parameters for the best fit are a rotational velocity of
$v_{max}=300$\kms, an optical depth $\tau=1\times10^7$ and temperature
of $T=1.5\times 10^{4}$K.
The optical depth and temperature can be translated into a column
density of $\log N_{HI} / \mathrm{atoms\ cm}^{-2} =
  20.5$.   
Our model is also able to constrain the angle
between the rotation axis and the observational line-of-sight to
$\theta=65^{\circ}\pm$. 
To determine the uncertainties we perform a Markov Chain Monte
Carlo calculation using an approximate analytic solution for the
Ly-$\alpha$ spectrum including rotational effects.

@@

Why is triple peaked emission not common in Lyman-alpha emitting
gaalaxies?   
The ubiquity of galaxy outflows overimposed to the rotation
transforms the line into the common double or single peaked
line.
Therefore, the absence of HI outflows in \tol\ is another special
feature that permitted a symmetric triple peaked lya line to emerge.

>>>>

\tol\ is a compact star forming dwarf galaxy that presents a
strong Ly-$\alpha$ emission \cite{Thuan97} with two important
features. 
First, it has symmetric profile  without any systemic displacement
from the expected line-center. 
Second, it shows a triple peaked profile. 
Commonly, the Ly-$\alpha$ line features a single or asymmetric double peak. 
These two special features in \tol\ had evaded a physical
interpretation so far.
We find that rotation and a multiphase ISM can independently explain
these two features.

Figure 1. summarizes our findings.
Dots represent the observational data for \tol with the
overplot from our best fit model from the full radiative transfer
simulation for a rotating sphere (continous line) and a clumpy ISM
(dashed line).

The parameters for the best fit are a rotational velocity of
$v_{max}=300$\kms, an optical depth $\tau=1\times10^7$ and temperature
of $T=1.5\times 10^{4}$K.
The optical depth and temperature can be translated into a column
density of $\log N_{HI} / \mathrm{atoms\ cm}^{-2} =
  20.5$.   
Our model is also able to constrain the angle
between the rotation axis and the observational line-of-sight to
$\theta=65^{\circ}\pm$. 
To determine the uncertainties we perform a Markov Chain Monte
Carlo calculation using an approximate analytic solution for the
Ly-$\alpha$ spectrum including rotational effects.

@@

%Why is triple peaked emission not common in Lyman-alpha emitting
%galaxies?   
%The ubiquity of galaxy outflows overimposed to the rotation
%transforms the line into the common double or single peaked
%line.
%Therefore, the absence of HI outflows in \tol\ is another special
%feature that permitted a symmetric triple peaked lya line to emerge.
